---
title: "Computing the variance of K12 under independent uniform marginals"
author: "Tuomas Rajala"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Bivariate K variance under CSR}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

How to compute the covariance of K12 estimator under independent binomial processes.

The estimator we work with is of type defined by Ward & Ferrandino (1), see also Wiegand et al. (2):

$$\hat K_{12}(r) = c(r)\sum_{x\in X_1}\sum_{y\in X_2}1_W(x)1_W(y)1(|x-y|<r)$$

where the constant is for edge correction and normalisation,

$$c(r)=\frac{\pi r^2 |W|^2}{n_1n_2 \Gamma_W(r)}$$

with 

$$\Gamma_W(r) = \int_W\int_W 1_{b(o,r)}(x-y)dxdy = 2\pi\int_0^r t \bar\gamma_W(t) dt$$
with isotropised set covariance $\bar\gamma_W$ of $W$.


```{r}
library(bivpower)
library(spatstat)
```

Setup example:

```{r}
# Window
W <- square(100)
V <- area(W)
# Abundances
n <- c(50, 100)
# Ranges for estimation
r <- 1:50
```

Computing parts of the variance requires MC integration. Let's do all bits in one go:

```{r, cache=TRUE}
# Ball-ball integral with formula for rectangles
I1 <- double_integral_ball(W, r)
# Triple integral, using MC. Average over few more efficient.
G3l <- sapply(1:5, function(...) triple_integral_ball_ball_mc(W, r) )
I4  <- rowMeans(G3l) 
a1 <- I1^2/V^4
a2 <- I4/V^3
a4 <- I1/V^2
#
K12norm <- area(W)^2 * pi * r^2 / ( prod(n) * I1)
# Compile the variance:
tvar <- K12norm^2 * ((sum(n)-2)*a2 + a4 - (sum(n)-1)*a1) * prod(n)

# 
# K12 Estimator, fixed n and W:
K12 <- function(x) K12norm * K12_doublesum(cbind(x$x, x$y), as.integer(x$marks), r)
```

Compare to empirical:

```{r, cache=TRUE}
nsim <- 5000
xl <- rmpoint(n, win = W, nsim = nsim)
k12 <- sapply(xl, K12)

# Empiricals
memp <- apply(k12, 1, mean)
vemp <- apply(k12, 1, var)

```


```{r, fig.width = 9, fig.height = 4}
par(mfrow=c(1,2))
plot(r, pi * r^2, main = "mean")
lines(r, memp)
plot(r, sqrt(tvar), main = "variance")
lines(r, sqrt(vemp) )
```

Some MC error (will go away by increasing "n" in `triple_integral_ball_ball_mc` and `nsim`)

## References

1. J.S. Ward and F.J. Ferrandino. New derivation reduces bias and increases power of Ripley’s L index. Ecological Modelling, 116(2-3):225–236, 1999.
2. T. Wiegand, P. Grabarnik, and D. Stoyan. Envelope tests for spatial point patterns with and without simulation. Ecosphere, 7(June):1–18, 2016.
