---
title: "Simulating the simple bivariate point pattern model with cross-interaction"
author: "Tuomas Rajala"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Bivariate Model}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

Very quick demo of how to simulate from the model we introduced in our paper.

```{r}
library(bivpower)
library(spatstat)
```

## Fixed abundances

The simulation is based on the ```lambda``` and  ```rcox``` functions and is essentially just a wrapper for these two. The wrapper was written for fixed abundance simulations, and it is quite easy to use: 

```{r}
# Window
W <- square(100)
# interaction
b <- -0.5
# abundances
n <- c(50, 80)
# scale
sig <- 10/2

# Simulation 
x <- sim_biv_model(n = n, W = W, theta = b, sigma = sig)

head(x)
```

To get directly a ```ppp``` -object use

```{r, fig.width=5, fig.height=5}
xppp <- sim_biv_model(n = n, W = W, theta = b, sigma = sig, asppp = TRUE)
plot(xppp, cols=1:2)
```

To provide fixed species 1

```{r,fig.width=5, fig.height=4}
x1 <- x[ x[,3] == 1, -3]
z <- sim_biv_model(n = n, W = W, theta = b, sigma = sig, x1 = x1, asppp=TRUE)
same1 <- listof(xym2ppp(x, W= W), z) # xym2ppp and ppp2xym provided
plot(same1, cols=1:2)
```


## Free abundance simulation

Two ways are shown here. First is direct and second is more illustrative.

### Poisson variation

Given 'n', samples $n_i\sim Poisson(n_i)$

```{r}
# Simulation 
xf <- sim_biv_model(n = n, W = W, theta = b, sigma = sig, free=TRUE)
table(xf[,3])
```

### Unconditional simulation by hand

As noted above, the wrapper consists of two stages. Let's go through them to see how to simulate unconditionally from the model.

```{r, fig.width=4.2, fig.height=5}
# 1.  Generate type 1 from Poisson(lam1, W)
lam1 <- 50/area(W)
n1 <- rpois(1, lam1*area(W))  # redundancy just to be clear
bbox <- with(W, cbind(xrange, yrange)) # bounding box
x1 <- apply(bbox, 2, function(ab) runif(n1, ab[1], ab[2]) ) # type 1 simulation

# 2. Create the lam2|x1. 
alpha0 <- log(80/area(W)) # would be equal to log(lambda) if interaction b = 0
lam2 <- lambda( x1,  kernel ="gauss", sigma = sig, type = "product", alpha = c(alpha0, b) )

print(lam2)


# 3. Simulate type 2 using thinning
x2 <- rcox(lambda = lam2, bbox = bbox) 

# Or fixed n using MH
x2f <- rcox(lambda = lam2, bbox = bbox, n = 50 , iter = 1e4)

# Done. Visualise the lam2

lam2im <- coxintensity2matrix(lam2, bbox = bbox, nx = 123)
plot(lam2im, col = gray.colors(12))
points(x1)
points(x2$x, pch=2, col=2)
```








